A Data-Driven Framework for Neural Field Modelling

### 2010-11-11 ltx/first_order_framework/LaTeX_Manuscript.tex

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the state and parameter estimation algorithm several experiments were performed. The results of the first experiment demonstrates the estimation performance using a Monte Carlo approach for a single set of parameters. The second experiment demonstrates the performance with a selection of different connectivity kernel parameters. The third experiment shows the dependance of the estimation accuracy on the parameters of the sigmoidal nonlinearity. The final experiment shows the effect of the width of the sensor kernel on the estimation performance.

@@

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good agreement with the actual values of the connectivity kernel parameters, however the time constant parameter, $\xi$, is outside the estimate distribution (with bias of 2.78\%).

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range ([-3.29,3.77]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(b) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(a)), but is missing the higher-frequency detail.

@@

To study the robustness of the estimation algorithm, the performance was studied for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Figure~\ref{fig:Figure8}. By changing the weights of the connectivity kernel we are effectively changing gain of the system and the range of the field amplitudes. Since the sigmoidal nonlinearity is remaining static, we are changing the operating region on the sigmoid. Therefore, we expect our estimation accuracy to vary.

Losing higher frequency peaks
Losing higher frequency troughs

For each subset of parameters the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95\% confidence interval is indicated by the grey shaded region. Parts a), b) and c) of Figure~\ref{fig:Figure8} show results from the Mexican hat kernel parameters were varied. In all cases the actual kernel fell within the confidence interval of the estimated kernel. The kernel in Part a) has the least amount of excitation and the most amount of inhibition. Therefore of the three Mexican hat kernels shown, this one will have the least amount of activation. The result is an overestimate of the central excitatory region of the kernel. Conversely, in Part c) of the figure the actual kernel has the most amount of excitation and the least amount of inhibition. Therefore

The mean kernel estimate for Part b) shows very good accordance with the actual kernel. In the other Parts a) and c) there is an over estimate and an underestimate of the kernel, respectively. These bias are a result

\parham{this bit should change as well:} 
For parts b) and c) the figure verifies that the shape of the estimated connectivity kernel is in good accordance with the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias in the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:Figure7}(c).

@@

Figure~\ref{fig:Figure11} shows a Hinton diagram demonstrating the estimation accuracy (or RMSE) of the connectivity kernel for various parameters of the sigmoid varying the slope, $\varsigma$, and the threshold, $v_0$, whilst the other parameters remain fixed. The kernel and time constant parameters used in this experiment were the one that gave the best results from the previous experiment. The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations. From the figure it can be seen that the estimation accuracy depends on the characteristics of the sigmoid function. When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However as the slope is increased or the cut-off is increased, more information is being cut off by the activation function. This is illustrated in figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the the bottom left corner through to near top right corner of Figure~\ref{fig:Figure11} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Results for Varying the Sensor Kernel Width}

@@

\tiny{\begin{tabular}{lcccccccccccc}
Experiment&\multicolumn{4}{c}{IV}&\multicolumn{4}{c}{V}&\multicolumn{4}{c}{VI}\\
 Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$\\
\hline\hline
Actual Value&$15$&$0$&$0$&$30$&$0$&$0$&$50$&$0$&$0$\\
Mean Estimate&$35.68$&$-18.16$&$0.81$&$0.923$&$40.43$&$-12.01$&$ 0.35$&$0.924$&$61.46$&$-15.03$&$0.27$&$0.924$\\
Std. Dev.&$16.72$&$11.77$&$0.53$&$0.002$&$11.47$&$7.8$&$0.36$&$0.003$&$11.51$&$7.56$&$0.29$&$0.003$\\
Bias (\%)&$137.87$&$-$&$-$&$2.56$&$34.77$&$-$&$-$&$2.67$&$22.92$&$-$&$-$&$2.67$\\
\end{tabular}}

>>>>

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
Experiments were performed to demonstrate the performance of the state and parameter estimation algorithm. The first experiment evaluates the estimation performance using a Monte Carlo approach for a single set of parameters. The second experiment evaluates the performance with a selection of different connectivity kernel parameters. The third experiment shows the dependance of the estimation accuracy on the parameters of the sigmoidal nonlinearity. The final experiment shows the effect of the width of the sensor kernel on the estimation performance.

@@

Histograms of the parameter estimates obtained over the 150 realisations are shown in Figure~\ref{fig:Figure7}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given with the results from the next Section in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good agreement with the actual values of the connectivity kernel parameters, however the time constant parameter, $\xi$, is outside the estimate distribution (with bias of 2.78\%).

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range \dean{([-3.29,3.77]~mV)}. The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(b) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(a)), but is missing the higher-frequency detail.

@@

To study the robustness of the estimation algorithm, the performance was studied for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Figure~\ref{fig:Figure8}. For Parts a), b) and c) a Mexican hat connectivity kernel was used. For each subset of parameters the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case the actual kernel was within the 95~\% confidence of kernel estimate. The statistics for the kernel parameters estimates for each of the parameter sets are shown in Table~\ref{tab:Parameters estimates}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. Parts d), e) and f) of Figure~\ref{fig:Figure8} show results when the kernel support in the estimator differed from the actual data. In these cases the data was generated using a single positive (excitatory) Gaussian kernel. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates, for the majority of the kernel space until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Importantly, although the reconstructed kernel estimate is in good accord with the actual kernel, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the support of the kernel basis functions are precisely known, the exact meaning of the parameters in isolation is not clear and the net reconstructed kernel should be considered in a practical setting.

@@

Figure~\ref{fig:Figure11} shows a Hinton diagram demonstrating the affect of the sigmoid parameters on the parameter estimation accuracy. The RMSE of the reconstructed connectivity kernel was calculated for various parameters of the sigmoid by varying the slope, $\varsigma$, and the threshold, $v_0$. The kernel and time constant parameters used in this experiment were the one that gave the best results from the previous experiment. The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). From the figure it can be seen that the estimation accuracy depends on the characteristics of the sigmoid function. When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more information is being cut off or attenuated by the activation function. This is illustrated in figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:Figure11} are plotted. This appears to have a detrimental effect on the estimation accuracy.

\subsubsection{Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for observation kernels of different width. To test the affect of the electrode kernel parameter on the estimation

@@

\tiny{\begin{tabular}{lcccccccccccc}
Experiment&\multicolumn{4}{c}{IV}&\multicolumn{4}{c}{V}&\multicolumn{4}{c}{VI}\\
 Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$\\
\hline\hline
Actual Value&$15$&$0$&$0$&$30$&$0$&$0$&$50$&$0$&$0$\\
Mean Estimate&$35.68$&$-18.16$&$0.81$&$0.923$&$40.43$&$-12.01$&$ 0.35$&$0.924$&$61.46$&$-15.03$&$0.27$&$0.924$\\
Std. Dev.&$16.72$&$11.77$&$0.53$&$0.002$&$11.47$&$7.8$&$0.36$&$0.003$&$11.51$&$7.56$&$0.29$&$0.003$\\
Bias (\%)&$137.87$&$-$&$-$&$2.56$&$34.77$&$-$&$-$&$2.67$&$22.92$&$-$&$-$&$2.67$\\
\end{tabular}\label{tab:GaussianKernelResults}}
